title: "Exploration Recocrop"
author: "Oscar Bautista"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ecocrop exploration

Load library

```{r}
library(Recocrop)
```

```{r}
crop.list <-ecocropPars()
```

Get parameters for a specific crop

```{r}
x <- ecocropPars("potato")
potato <- x$parameters
```

```{r}
m <- ecocrop()
crop(m) <- potato
```

## Plot crop parameters

You can also embed plots, for example:

```{r m}
plot(m)
```

Explore the default parameters for each crop

```{r}
fname <- system.file("parameters/ecocrop.rds", package="Recocrop")
d <- readRDS(fname)
head(d)
```

Add parameter

```{r}
crop(m) <- cbind(clay=c(0,0,10,20))
```

Plot parameters and verify the new parameter (Clay) was added

```{r}
plot(m)
```

add predictors to tavg variable

```{r}
dp <- cbind(tavg=c(10,12,14,16,18,20,22,20,18,16,14,12), prec=seq(50,182,12))
dynamicPredictors(m) <- dp
staticPredictors(m) <- cbind(clay=12)
```

Run model with modified parameters, set labels and axes

```{r}

x <- run(m)
x
y <- matrix(round(x, 1), nrow=2)
colnames(y) <- month.abb
rownames(y) <- c("day1", "day15")
y
dates <- as.Date(paste0("2000-", rep(1:12, each=2), "-", rep(c(1,15), 12)))
```

Plot the suitability curve

```{r, figures-side}
plot(dates, x, las=1, ylab="suitability", xlab="")
lines(dates, x, col="red")

```

## Predictors

load the terra library to include the spatial dimension

```{r}
library(terra)
```

explore the default spatial parameters in the model

```{r}
fta <- system.file("ex/ta.tif", package="Recocrop")
fpr <- system.file("ex/pr.tif", package="Recocrop")
fph <- system.file("ex/ph.tif", package="Recocrop")
# monthly average temperature
ta <- rast(fta)
# monthly precipitation
pr <- rast(fpr)
# pH
ph <- rast(fph)
# just for plotting
preds <- c(mean(ta), sum(pr), ph)
names(preds) <- c("tavg", "prec", "ph")
plot(preds)
```

```{r}
## make ecocrop model
crop <- ecocropPars("maize")
m <- ecocrop(crop)
control(m, get_max=TRUE)
# run the model, make sure to match
# the predictor and parameter names
mz <- predict(m, tavg=ta, prec=pr, ph=ph, wopt=list(names="maize"))
plot(mz)
```

```{r}
# output by 15 day period
control(m, get_max=FALSE)
mz12 <- predict(m, tavg=ta, prec=pr, ph=ph)
plot(mz12[[14:15]])

```

```{r}
control(m, lim_fact=TRUE)
mzlim <- predict(m, tavg=ta, prec=pr+50, ph=ph)
plot(mzlim[[14]])
control(m, get_max=TRUE)
```

Modification of pH parameter

```{r}
# make the soil more acidic
ph2 <- ph - 1.25
control(m, get_max=TRUE)
mzph <- predict(m, tavg=ta, prec=pr, ph=ph2)
plot(mzph)
```

Modify rainfall by adding 30 units/month

```{r}
# more rainfall
control(m, get_max=TRUE)
pr2 <- pr + 30 # for each month
mzpr <- predict(m, tavg=ta, prec=pr2, ph=ph)
plot(mzpr)
```

Plot base and modified models (pH and rainfall)

```{r}
s <- c(mz, mzph, mzpr)
names(s) <- c("base", "ph", "prec")
plot(s)
```

Parallel plot for two or more crop models

```{r}

crop <- ecocropPars("pearl millet")
m <- ecocrop(crop)
control(m, get_max=TRUE)
pm <- predict(m, prec=pr, tavg=ta, wopt=list(names="Pearl millet"))

crop1name <- "Bean, Common"
crop1n <- "Bean"
crop1 <- ecocropPars(crop1name)
mod.crop1 <- ecocrop(crop1)

#get max value in data series
control(mod.crop1, get_max=TRUE)
c1 <- predict(mod.crop1, prec=pr, tavg=ta, wopt=list(names=crop1n))


crops <- c(mz, pm, c1)
plot(crops)
```

Interactive map

```{r}
library(leaflet)
library(raster)
```

Set visualization colors

```{r}
pal <- colorNumeric(c("green", "yellow", "red"), values(mz),
  na.color = "transparent")
palmzph <- colorNumeric(c("green", "yellow", "red"), values(mzph),
  na.color = "transparent")

```

Add the maize models (Control and acidic) to an interactive leaflet map

```{r}
mz.lay <- raster(mz)
mzph.lay <-raster(mzph)
leaflet() %>% 
  addTiles() %>%
  addRasterImage(mz.lay, colors = pal, opacity = 1, 
                 group= "control" ) %>%
  addRasterImage(mzph.lay, colors = palmzph, opacity = 1, 
                 group= "acidic" ) %>%
  addLegend(values = seq(from = 0, to = 1, by= 0.1),
            pal = pal,
  title = "maize") %>%
  addLayersControl(overlayGroups = c("control", "acidic"),
    options = layersControlOptions(collapsed = FALSE))
```
